Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
refs

Recent improvements in instrumentation have revolutionized the kind of
kinematic studies that can be performed on galaxies with small angular
sizes on the sky.
For instance, spatially resolved spectra on dwarf galaxies show signs
of coherent rotation \citep{2009A&A...493..871S}.
Some Compact Dwarf Galaxies actually show gas kinematics that range
from pure rotation to high velocity dispersion without a clear
rotation pattern.
\citep{2015A&A...577A..21C,2017A&A...600A.125C}
Some of those star forming galaxies are expected to have high neutral
gas contents and low dust contents and thus show \lya\ emission with
rotation imprints
\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}

The computational \lya modeling of systems showing rotation has only
been studied until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morpholgy.
They found that rotation indeed introduces changes, the most
noticeable change is the dependence of the spectra with the viewing
angle with respect to the rotation axis.

@@

Current observational facilities have the capability of spatially
resolving the extent of a LAE.
For instance \cite{Prescott14} presented observational results of a
Doppler shift when taking spectra at two oposite sides of a LAE.

@@

\subsubsection{Relationship to MUSE observations}

In order to give the reader a feeling of the ranges of observational
data that can be explained by the models presented here we perform a
comparison against observational data from MUSE-Wide survey (MWS)
\citep{2017A&A...606A..12H}.

MWS provides public access to 831 spectra out of which 237 contain a
detected \lya line as dominante. 
The public dataset reports as the redshift of the galaxy the position
of the dominant peak.
We use this redshift to convert the spectra to rest-frame and compute
the standard deviation and skewness statitstics. 
Given the uncertainty on the galaxy redshift we do not compute the
valley/peak ratio. 
We do not try to perform a full fit to all the spectra but only to
explore whether the region of parameter space explored in our RT
simulations shows any crossing with the results from the MWS.

@@

Our work confirms that rotation produces features in the \lya line
strong enough that can be expected to be detected in observations and
that should be taken into account into \lya modelling; the Doppler
Boost offers an easy-to-implement approximation to explore such
influence into already existing RT simulations.

>>>>

Recent advances in instrumentation have revolutionized the kind of
kinematic studies that can be performed on galaxies with small angular
sizes on the sky.
For instances, spatially resolved spectra on compact dwarf galaxies
are now possible, and they confirm signs of gas kinematics showing
pure rotation kinematics \citep{2015A&A...577A..21C,2017A&A...600A.125C}
Recent mapping of high redshift circumgalactic regions are also able
to resolve kinematic evicend of large scale rotation
\citep{2018MNRAS.473.3907A} 
These systems with star formation, some neutral gas and low dust
contents are the kind of systems that are naturally expected to have
\lya emission.
What is the expected imprint of rotation on a resonant line such as
the \lya line?

The computational \lya modeling of rotating systems has only
been studied until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morpholgy.
They found that rotation indeed introduces changes, the most
noticeable change is the dependence of the spectra with the viewing
angle with respect to the rotation axis.

@@

Current observational facilities have the capability of spatially
resolving the extent of a LAE.
For instance \cite{Prescott14} presented observational results of a
Doppler shift when taking spectra at two oposite sides of a large
($\approx 80$kpc) LAE.

@@

\subsubsection{Comparison against the MUSE-Wide Survey}

In order to give the reader a feeling of the ranges of observational
data that could be explained by the models presented here, 
we perform a comparison against observational data from the MUSE-Wide
Survey (MWS) \citep{2017A&A...606A..12H}.

MWS provides public access to 831 spectra out of which 237 contain a
detected \lya line as dominant. 
The public dataset reports as the redshift of the galaxy the position
of the dominant peak.
We use this redshift to convert the spectra to a rest-frame and compute
the standard deviation and skewness statitstics. 
Given the uncertainty on the galaxy redshift we do not compute the
valley/peak ratio. 
We do not try to perform a full fit to all the spectra
\citep[e.g.][]{2017A&A...608A.139G} and wish only to
know whether the region of parameter space explored in our RT
simulations shows any crossing with the results from the MWS or
shows any sign of atypical features.

Figure \ref{fig:muse} shows the results of this experiment. 
The small white circles correspond to the 217 \lya emitters in the MWS
observations; the large black circles to the results of the
simulations presented in this paper. 
This shows that the systems modeled in this paper actually present an
overlap with statistics derived from observations. 
The region with standard deviation $>200$\kms could also be explained
by systems with larger optical depths than we have used here. 
However, the systems with a positive skewness, that is a line
unbalanced towards the blue could be the sign of inflowing
kinematics.

@@

Our work confirms that rotation produces features in the \lya line
strong enough that can be expected to be detected in observations and
that should be taken into account into \lya modelling; the Doppler
Boost offers an easy-to-implement approximation to explore such
influence into already existing RT simulations and provide stronger
footing to the interpretation of \lya kinematics maps
\citep[e.g][]{2018MNRAS.473.3907A}.
